---
title: "Winter Sampling"
output: html_notebook
---



```{r}
library(tidyverse)
library(readxl)
library(sf)
library(ggrepel)
```

First read in costline vectors

```{r}
coastlines <- read_sf("../genoscape_maps/shapefiles/ne_shapefiles/ne_10m_coastline.shp") %>% 
  st_set_crs(st_crs(4326))

```
Now, let's just crop out the part that we want.  This is somewhat key:
right here we will define the extent in lat-long of the region that
we want to plot:
```{r}

domain <- c(
  xmin = -130, 
  xmax = -65,
  ymin = 10,
  ymax = 60
)

coast_cropped <- st_crop(coastlines, domain)

```

Then plot the cropped part:
```{r}
ggplot(coast_cropped) + 
  geom_sf() +
  coord_sf()
```

### All the lines

OK, that was pretty reasonable.  Now crop all the lines to `domain` and plot
them, and put the genoscape on top of that, too.
```{r}
countries_cropped <- read_sf("../genoscape_maps/shapefiles/ne_shapefiles/ne_10m_admin_0_boundary_lines_land.shp") %>%
  st_crop(domain) %>% 
  st_set_crs(st_crs(4326))

states_cropped <- read_sf("../genoscape_maps/shapefiles/ne_shapefiles/ne_10m_admin_1_states_provinces_lines.shp") %>%
  st_crop(domain) %>% 
  st_set_crs(st_crs(4326))

nonbreeding_range <- read_sf("../genoscape_maps/shapefiles/LOSH/logshr_range_2021.gpkg") %>%
  filter(season == "nonbreeding")

```

Now, plot it all.  Notice we do it by just adding layers.
```{r}
mapg <- ggplot() +
  geom_sf(data = coast_cropped, fill = "grey97", color = "grey") +
  geom_sf(data = countries_cropped, fill = "grey97", color = "grey") +
  geom_sf(data = states_cropped, fill = "grey97", color = "grey") +
  geom_sf(data = nonbreeding_range, fill = "lightsteelblue2", lwd = 0 , alpha = 0.7) +
  theme_bw() 

mapg +
  coord_sf() 
```

Now for sample locations

## Import metadata
```{r}
meta <- read.csv("../meta/NewLoshSamples.csv")
database <- read.csv("../meta/database_corrected.csv")
database_ubreeding <- database %>% 
  filter(Breeding == 'U')
database
```

## Any transformations
```{r}
winter_months_relaxed <- c("NOV", "DEC", "JAN", "FEB")

database <- database %>%
  mutate(CollectionDate = as.Date(CollectionDate, format = "%m/%d/%Y"),
  Month = toupper(format(CollectionDate, "%b")))

database_winter <- database %>% 
  filter(Breeding == 'N', Sequenced == 'N') %>%
  #estimate lat longs
  mutate(Lat = if_else(CityTown == "Myrtle Beach" & is.na(Lat), 33.6954, Lat),
         Long = if_else(CityTown == "Myrtle Beach" & is.na(Long), -78.8802, Long)) %>% 
  mutate(Lat = if_else(CityTown == "Conway" & is.na(Lat), 33.8360, Lat),
         Long = if_else(CityTown == "Conway" & is.na(Long), -79.0478, Long)) %>% 
  mutate(Lat = if_else(State == "WV" & is.na(Lat), 37.6, Lat),
         Long = if_else(State == "WV" & is.na(Long), -80.5, Long)) %>% 
  filter(Month %in% winter_months_relaxed)

meta_winter <- meta %>%
  filter(Month %in% winter_months_relaxed)

```

## Import Chabot 03.11.24 Feathers
```{r}
amy_winter_database <- read.csv("../meta/Amy_LOSH_Feathers_SentMar24_SortedByHFJul24.csv") %>%
  filter(Retrix_Calamus_Num >= 1, Retrix_Calamus_Num != "Missing") %>%
  mutate_at(vars(Latitude, Longitude), as.numeric) %>% 
  mutate(Longitude = if_else(Longitude > 0, -Longitude, Longitude)) %>% 
  mutate(Longitude = if_else(State == "Jalisco", -103.3473, Longitude)) %>%  # coordinates for guadalajara
  mutate(Longitude = if_else(State == "Michoacan", -101.9900, Longitude)) %>% # coordinates for michoacan centroid
  mutate(Longitude = if_else(State == "Aguascalientes", -102.2916, Longitude)) %>% # coordinates aguascalientes
  mutate(Sample.Date = as.Date(Sample.Date, format = "%Y/%m/%d"),
         Month = toupper(format(Sample.Date, "%b"))) %>% 
  filter(Month %in% winter_months_relaxed | Season == "Wintering")
  
```


```{r}

mapg <- ggplot() +
  geom_sf(data = coast_cropped, fill = "grey97", color = "grey") +
  geom_sf(data = countries_cropped, fill = "grey97", color = "grey") +
  geom_sf(data = states_cropped, fill = "grey97", color = "grey") +
  geom_sf(data = nonbreeding_range, fill = "lightsteelblue2", lwd = 0 , alpha = 0.8) +
  theme_bw() +
  xlab("Longitude") +
  ylab("Latitude") +
  xlim(-130, -65) + 
  ylim(10, 60) +
  geom_point(data = meta_winter,
             aes(x = Long, y = Lat),
             color = "black",
             size = 2.5, alpha = 0.5) +
  geom_point(data = database_winter,
             aes(x = Long, y = Lat),
             color = "black",
             size = 2.5, alpha = 0.5) +
  geom_point(data = amy_winter_database,
             aes(x = Longitude, y = Latitude),
             color = "black",
             size = 2.5, alpha = 0.5) +
  geom_point(data = subset(MLZ_winter, state_prov == "San Luis Potosi"),
             aes(x = dec_long, y = dec_lat),
             color = "green",
             size = 2.5, alpha = 0.3) +
  geom_point(data = subset(MLZ_winter, state_prov == "Zacatecas"),
             aes(x = dec_long, y = dec_lat),
             color = "green",
             size = 2.5, alpha = 0.3) +
  geom_point(data = subset(MLZ_winter, state_prov == "Sinaloa"),
             aes(x = dec_long, y = dec_lat),
             color = "green",
             size = 2.5, alpha = 0.3) +
  geom_point(data = subset(MLZ_winter, state_prov == "Guanajuato"),
             aes(x = dec_long, y = dec_lat),
             color = "green",
             size = 2.5, alpha = 0.3) +
  geom_point(data = subset(MLZ_winter, state_prov == "Morelos"),
             aes(x = dec_long, y = dec_lat),
             color = "green",
             size = 2.5, alpha = 0.3) +
  geom_point(data = subset(MLZ_winter, state_prov == "Queretaro"),
             aes(x = dec_long, y = dec_lat),
             color = "red",
             size = 2.5, alpha = 0.3) +
  coord_sf()

mapg
#ggsave("losh_winter_database_plusMLZ_0724.svg", plot = mapg)
```
